% Copyright (c) 2016 The ALF project.
% This is the ALF project documentation.
% The ALF project documentation by the ALF contributors is licensed
% under a Creative Commons Attribution-ShareAlike 4.0 International License.
% For the licensing details of the documentation see license.CCBYSA.

\documentclass[10pt,Arial]{scrartcl}
\usepackage{graphicx}
\usepackage[margin=2.5cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{bbm}            
\usepackage[fleqn]{amsmath} 
\usepackage{amssymb}
\usepackage{wasysym}
\usepackage{color}
\usepackage{soul}
\usepackage{xspace}
\usepackage{bm}
\usepackage{subfigure}
\usepackage{tcolorbox}
\usepackage{dsfont}
\usepackage{footnote}
\usepackage{alltt}
\usepackage{xcolor}
\usepackage{lmodern}
\usepackage{listings}
\usepackage{url}
\usepackage{booktabs}
\usepackage{hyperref} 

\lstset{language=Fortran,
  basicstyle=\ttfamily,
  keywordstyle=\color{red},
  commentstyle=\color{blue},
  morecomment=[l]{!\ }% Comment only with space after !
  breakatwhitespace=false,
  keepspaces=true,
  showstringspaces=false,
  columns=flexible
}

\lstdefinestyle{custombash}{
language=bash,
  basicstyle=\ttfamily,
  keywordstyle=\color{red},
  commentstyle=\color{blue},
  morecomment=[l]{\#\ }% Comment only with space after #
  breakatwhitespace=false,
  keepspaces=true,
  showstringspaces=false,
  columns=flexible
}

\def\Tr{\mathop{\mathrm{Tr}}}
\def\Trf{\mathop{\mathrm{Tr}_{\mathrm{F}}}}
\makesavenoteenv{tabular}
\makesavenoteenv{table}
\setkomafont{author}{\large}
\setkomafont{date}{\large}

\newcommand{\mycomment}[1]{{\color{blue} #1}}
\newcommand{\FAcomment}[1]{{\color{red} #1}}

\makeindex
\begin{document}
%---------------------------------------------------------------------------------------------------------
\title{The \textit{ALF} (\textit{A}lgorithms for \textit{L}attice \textit{F}ermions) project release 0.5}
\subtitle{Documentation for the  auxiliary field quantum Monte Carlo  code.}
\author{Martin Bercx,  Florian Goth,  Johannes S. Hofmann, Fakher F. Assaad }
%---------------------------------------------------------------------------------------------------------
\maketitle

Copyright \textcopyright ~2016, The \textit{ALF} Project.\\
This is the ALF Project Documentation by the ALF contributors.
It is licensed under a Creative Commons Attribution-ShareAlike 4.0 International License.
You are free to share and benefit from this documentation as long as this license is preserved
and proper attribution to the authors is given. For details see the ALF project
homepage \url{alf.physik.uni-wuerzburg.de}.
\tableofcontents
\clearpage
\input{intro}
\input{model}
\input{implementation}
\input{files}
\input{walkthrough_hubbard}
\input{other_models}
\input{analysis}
\input{running}
\input{performance}
\input{conclusion}
\addcontentsline{toc}{section}{Acknowledgments}
\input{acknowledgment}
\addcontentsline{toc}{section}{References}
%\bibliographystyle{./prXsty} 
%\bibliography{./fassaad}

\begin{thebibliography}{10}

\bibitem{Blankenbecler81}
R. Blankenbecler, D.~J. Scalapino, and R.~L. Sugar, Phys. Rev. D {\bf 24},
  2278  (1981).

\bibitem{Assaad08_rev}
F. Assaad and H. Evertz,  in {\em Computational Many-Particle Physics},
  Vol.~739 of {\em Lecture Notes in Physics}, edited by H. Fehske, R.
  Schneider, and A. Wei{\ss}e (Springer, Berlin Heidelberg, 2008), pp.\
  277--356.

\bibitem{Wu04}
C. Wu and S.-C. Zhang, Phys. Rev. B {\bf 71},  155115  (2005).

\bibitem{Wei16}
Z.~C. Wei, C. Wu, Y. Li, S. Zhang, and T. Xiang, Phys. Rev. Lett. {\bf 116},
  250601  (2016).

\bibitem{White89}
S. White, D. Scalapino, R. Sugar, E. Loh, J. Gubernatis, and R. Scalettar,
  Phys. Rev. B {\bf 40},  506  (1989).

\bibitem{Sugiyama86}
G. Sugiyama and S. Koonin, Annals of Physics {\bf 168},  1   (1986).

\bibitem{Sorella89}
S. Sorella, S. Baroni, R. Car, and M. Parrinello, EPL (Europhysics Letters)
  {\bf 8},  663  (1989).

\bibitem{Milat04}
I. Milat, F. Assaad, and M. Sigrist, Eur. Phys. J. B {\bf 38},  571  (2004),
  http://xxx.lanl.gov/cond-mat/0312450.

\bibitem{Bercx09}
M. Bercx, T.~C. Lang, and F.~F. Assaad, Phys. Rev. B {\bf 80},  045412  (2009).

\bibitem{Schattner15}
Y. Schattner, S. Lederer, S.~A. Kivelson, and E. Berg, Phys. Rev. X {\bf 6},
  031028  (2016).

\bibitem{Xu16}
X.~Y. {Xu}, K.~S.~D. {Beach}, K. {Sun}, F.~F. {Assaad}, and Z.~Y. {Meng},
  ArXiv:1602.07150  (2016).

\bibitem{Assaad16}
F.~F. Assaad and T. Grover, Phys. Rev. X {\bf 6},  041049  (2016).

\bibitem{Assaad99a}
F.~F. Assaad, Phys. Rev. Lett. {\bf 83},  796  (1999).

\bibitem{Capponi00}
S. Capponi and F.~F. Assaad, Phys. Rev. B {\bf 63},  155114  (2001).

\bibitem{Beach04}
K.~S.~D. Beach, P.~A. Lee, and P. Monthoux, Phys. Rev. Lett. {\bf 92},  026401
  (2004).

\bibitem{Assaad04}
F.~F. Assaad, Phys. Rev. B {\bf 71},  075103  (2005).

\bibitem{Lang13}
T.~C. Lang, Z.~Y. Meng, A. Muramatsu, S. Wessel, and F.~F. Assaad, Phys. Rev.
  Lett. {\bf 111},  066401  (2013).

\bibitem{Li15}
Z.-X. Li, Y.-F. Jiang, and H. Yao, New Journal of Physics {\bf 17},  085003
  (2015).

\bibitem{Broecker14}
P. Broecker and S. Trebst, Journal of Statistical Mechanics: Theory and
  Experiment {\bf 2014},  P08015  (2014).

\bibitem{Assaad14}
F.~F. Assaad, Nat Phys {\bf 10},  905  (2014).

\bibitem{Assaad13a}
F.~F. Assaad, T.~C. Lang, and F. Parisen~Toldin, Phys. Rev. B {\bf 89},  125121
   (2014).

\bibitem{Assaad15}
F.~F. Assaad, Phys. Rev. B {\bf 91},  125146  (2015).

\bibitem{Xu16a}
X.~Y. Xu, Y. Qi, J. Liu, L. Fu, and Z.~Y. Meng, arXiv:1612.03804  (2016).

\bibitem{Hohenadler14}
M. Hohenadler, F. Parisen~Toldin, I.~F. Herbut, and F.~F. Assaad, Phys. Rev. B
  {\bf 90},  085146  (2014).

\bibitem{Ulybyshev2013}
M.~V. Ulybyshev, P.~V. Buividovich, M.~I. Katsnelson, and M.~I. Polikarpov,
  Phys. Rev. Lett. {\bf 111},  056801  (2013).

\bibitem{Brower12}
R. Brower, C. Rebbi, and D. Schaich, PoS(Lattice 2011)056 (arXiv:1204.5424)  .

\bibitem{Jureca16}
{J\"ulich Supercomputing Centre}, Journal of large-scale research facilities
  {\bf 2},  A62  (2016).

\end{thebibliography}
\addcontentsline{toc}{section}{License}
\input{license}

\end{document}
